First Season MWA EoR Power Spectrum Results at Redshift 7

### 2016-06-29 ms.tex
EL comments

\subsection{Preprocessing}\label{subsec:preprocessing}
The preprocessing step converts the data from the non-standard memory dump format 
from the correlator's GPUs to a standard UVFITS format \citep{Greisen:2012}, 
phases the data to the center of the pointing,
performs 
flagging, averages in time and freqeuncy, and writes to disk. For this step we use the 
\cotter package \citep{Offringa:2015}, which in turn calls the \texttt{AOFLAGGER}\footnote{\url{http://
aoflagger.sourceforge.net/}} to perform RFI flagging \citep{Offringa:2010}.

@@

Calibration is accomplished within the FHD package in two steps. First we match the raw 
data to model visibilities formed using a realistic simulation of the array response to our 
foreground model. This step results in independent complex gains for each antenna, 
frequency channel, and polarization. In the second step we impose restrictions on these 
gains motivated by our understanding of the spectral response of the instrument. By 
reducing the number of free parameters in this step we increase our signal to noise on 
calibration solutions and avoid absorbing unmodeled confusion source flux density into our gain 
solutions which has been shown to contaminate the EoR window \citep{Barry:2016}. 
Recently, \citealt{Patil:2016} demonstrated excess noise and a suppression of diffuse
foregrounds in LOFAR images after calibration and subtraction of bright discrete
foreground sources. They suggested a multi-frequency calibration solution as a potential
solution to these systematics, similar to the scheme implemented here. Indeed we have
not seen evidence of these systematics through simulations and direct propagation of
our noise \citep[Hazelton et al. 2016, in prep.]{Barry:2016}.

In the first calibration step we allow our complex gains to account for any direction-independent 
response on a per antenna, per frequency channel, per polarization basis. To 
define our gains in this way, we have made two simplifications. First we push all 
direction-dependence of the antenna response into the model of the primary beam. FHD is capable 
of using separate models for each antenna, though this has not been implemented in this 
work. The second simplification is to ignore any cross 
terms between our different axes, for example a mutual coupling between two antennas. In 
principle these terms would introduce baseline dependent gains, rather than antenna 
dependent. We have not yet seen evidence of these terms in MWA data, and so we currently 
neglect them in our solutions.

@@

After calibration, we form snapshot image cubes (frequency mapping to line-of-sight 
direction). The visibilities are gridded using the primary beam as the gridding kernel, placing 
the data in the holographic frame, which has been shown to be the optimal frame for 
combining images in the sense that it preserves all information for parameter estimation 
\citep{Morales:2009,Bhatnagar:2008}. We use a simulation-based primary beam model for MWA 
antennas developed by \citealt{Sutinjo:2015}. At this stage we also average in frequency by 
a factor of two by gridding pairs of frequency channels to the same $(u,v)$ plane, resulting in 
a frequency resolution of 160 kHz.

@@

The diffuse model used for this work is shown in Figure~\ref{fig:diffuse}. While the actual 
model image is uniform weighted with resolution $\sim\!6$~arcminutes, we show it 
smoothed to degree scales to emphasize the large scale structure and approximate what 
the MWA instrument observes with a natural weight.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{diffuse_sky.pdf}
\caption[Diffuse model used for subtraction]{
The diffuse foreground model within the EoR0 field used for foreground subtraction. This 
model was created using residual images from the golden data set. The image shown is 
smoothed to degree scales to emphasize the large scale structure and approximate an 
MWA snapshot natural weighting.
\label{fig:diffuse}
}
\end{center}
\end{figure}

@@

\section{Discussion}\label{sec:discussion}
Inspecting our 32 hour integrated power spectra we see two places where we are limited. 
First there is a region that is leakage-dominated at low $k$. The cause of this leakage is yet 
unknown, but likely due to imperfect calibration, beam models, and/or foreground models.
In order to improve on our limit in this regime, future analysis will require calibration which 
better accounts for the instrumental response, perhaps by increasing signal to noise 
through multiple-snapshot solutions or refined antenna response models. 
Through simulation, \citealt{Thyagarajan:2016} recently demonstrated the necessity of
 precise foregrounds and instrumental models in order to retain a clean EoR window.
Studies of the 
beam are underway \citep[e.g.][]{Neben:2015}, and will likely help to further isolate the 
foreground contaminates in our power spectra. The foreground model is a constant area of 
investigation, but improved spectral dependence in both the point source catalog and the 
diffuse model, as well as the addition of an all-sky Galactic model, will continue to improve 
the foreground subtraction and further unlock the EoR window.

@@

Finally, we place our best upper limits in context with the 21cm EoR field as a whole. A 
direct comparison between measurements from different instruments is difficult due to the 
varying methods, redshifts, and scales probed. However, if we approximate the power 
spectrum in units of mK$^2$ to be roughly flat over the scales probed by the MWA, PAPER, 
and GMRT, we can glean a view of the current state of the field. The results from this 
analysis are a modest improvement over the previous best MWA results \citep{Dillon:2015}, 
which is not surprising given that this integration contains about ten times more data, but 
is systematic limited. Our results are also consistent with the 
40 hour GMRT limit of $\Delta^2\le6.15\times10^4$~mK$^2$ at $k=0.5\, h$~Mpc$^{-1}$ and $z=8.6$ \citep{Paciga:2013}, though 
probe significantly different periods of the EoR. The frontrunner in terms of an upper limit is 
the PAPER experiment \citep{Ali:2015, Jacobs:2015, Parsons:2014}. The PAPER-64 limit of 
$\Delta^2\le502$~mK$^2$ at $z=8.4$ from a 135 day observing campaign is about one and a half orders of magnitude lower 
than our best limit in power units, though again probing a significantly different redshift. 
The MWA analysis is rapidly improving as we uncover and mitigate systematics to realize
the full sensitivity of the telescope.
The red lines in Figure~\ref{fig:1d_ps} indicate the noise level possible 
to reach with these data if systematics can be overcome.

>>>>

\subsection{Preprocessing}\label{subsec:preprocessing}
The preprocessing step converts the data from the non-standard memory dump format 
from the correlator's GPUs to a standard UVFITS format \citep{Greisen:2012}, 
phases the data to the center of the pointing,
performs 
flagging, averages in time and frequency, and writes to disk. For this step we use the 
\cotter package \citep{Offringa:2015}, which in turn calls the \texttt{AOFLAGGER}\footnote{\url{http://
aoflagger.sourceforge.net/}} to perform RFI flagging \citep{Offringa:2010}.

@@

Calibration is accomplished within the FHD package in two steps. First we match the raw 
data to model visibilities formed using a realistic simulation of the array response to our 
foreground model. This step results in independent complex gains for each antenna, 
80~kHz frequency channel, and polarization. In the second step we impose restrictions on these 
gains motivated by our understanding of the spectral response of the instrument. By 
reducing the number of free parameters in this step we increase our signal to noise on 
calibration solutions and avoid absorbing unmodeled confusion source flux density into our gain 
solutions which has been shown to contaminate the EoR window \citep{Barry:2016}. 
Recently, \citealt{Patil:2016} demonstrated excess noise and a suppression of diffuse
foregrounds in LOFAR images after calibration and subtraction of bright discrete
foreground sources. They suggested a multi-frequency calibration solution as a potential
solution to these systematics, similar to the scheme implemented here. Indeed we have
not seen evidence of these systematics through simulations and direct propagation of
our noise \citep[Hazelton et al. 2016, in prep.]{Barry:2016}.

In the first calibration step we allow our complex gains to account for any direction-independent 
response on a per antenna, per 80~kHz frequency channel, per polarization basis. To 
define our gains in this way, we have made two simplifications. First we push all 
direction-dependence of the antenna response into the model of the primary beam. FHD is capable 
of using separate models for each antenna, though this has not been implemented in this 
work. The second simplification is to ignore any cross 
terms between our different axes, for example a mutual coupling between two antennas. In 
principle these terms would introduce baseline dependent gains, rather than antenna 
dependent. We have not yet seen evidence of these terms in MWA data, and so we currently 
neglect them in our solutions.

@@

After calibration, we form snapshot image cubes (frequency mapping to line-of-sight 
direction). The visibilities are gridded using the primary beam as the gridding kernel, placing 
the data in the holographic frame, which has been shown to be the optimal frame for 
combining images in the sense that it preserves all information for parameter estimation 
\citep{Morales:2009,Bhatnagar:2008}. We use a simulation-based primary beam model for MWA 
antennas developed by \citealt{Sutinjo:2015} 
which incorporates an average embedded element pattern for the dipoles, as well as mutual
coupling between dipoles within an antenna.
At this stage we also average in frequency by 
a factor of two by gridding pairs of frequency channels to the same $(u,v)$ plane, resulting in 
a frequency resolution of 160 kHz.

@@

The diffuse model used for this work is shown in Figure~\ref{fig:diffuse}. While the actual 
model image is uniform weighted with resolution $\sim\!6$~arcminutes, we show it 
smoothed to degree scales to emphasize the large scale structure and approximate what 
the MWA instrument observes with a natural weighting.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{diffuse_sky.pdf}
\caption[Diffuse model used for subtraction]{
The diffuse foreground model within the EoR0 field used for foreground subtraction. This 
model was created using residual images from the golden data set. The image shown is 
smoothed to degree scales to emphasize the large scale structure and approximate an 
MWA snapshot natural weighting. Note that the negative brightness values are a 
consequence of not sampling the zero-spacing part of the $(u,v)$ plane.
\label{fig:diffuse}
}
\end{center}
\end{figure}

@@

\section{Discussion}\label{sec:discussion}
Inspecting our 32 hour integrated power spectra we see two places where we are limited. 
First there is a region that is leakage-dominated at low $k$. The cause of this leakage is yet 
unknown, but likely due to imperfect calibration, beam models, and/or foreground models.
In order to improve on our limit in this regime, future analysis will require calibration which 
better accounts for the instrumental response, perhaps by increasing signal to noise 
through multiple-snapshot solutions or refined antenna response models. 
Through simulation, \citealt{Thyagarajan:2016} recently demonstrated the necessity of
 precise foregrounds and instrumental models in order to retain a clean EoR window.
Studies of the 
beam are underway \citep[e.g.][]{Neben:2015, Sutinjo:2015}, and will likely help to further isolate the 
foreground contaminates in our power spectra. The foreground model is a constant area of 
investigation, but improved spectral dependence in both the point source catalog and the 
diffuse model, as well as the addition of an all-sky Galactic model, will continue to improve 
the foreground subtraction and further unlock the EoR window.

@@

Finally, we place our best upper limits in context with the 21cm EoR field as a whole. A 
direct comparison between measurements from different instruments is difficult due to the 
varying methods, redshifts, and scales probed. However, if we approximate the power 
spectrum in units of mK$^2$ to be roughly flat over the scales probed by the MWA, PAPER, 
and GMRT, we can glean a view of the current state of the field. 
The best result from this analysis is an upper limit on the power spectrum of 
$\Delta^2 \le 2.7 \times 10^4$~mK$^2$ at $k=0.27\, h$~Mpc$^{-1}$ and $z=7.1$.
This is a modest improvement over the previous best MWA results \citep{Dillon:2015}, 
which is not surprising given that this integration contains about ten times more data, but 
is systematic limited. Our results are also consistent with the 
40 hour GMRT limit of $\Delta^2\le6.15\times10^4$~mK$^2$ at $k=0.5\, h$~Mpc$^{-1}$ and $z=8.6$ \citep{Paciga:2013}, though 
probe significantly different periods of the EoR. The frontrunner in terms of an upper limit is 
the PAPER experiment \citep{Ali:2015, Jacobs:2015, Parsons:2014}. The PAPER-64 limit of 
$\Delta^2\le502$~mK$^2$ at $z=8.4$ from a 135 day observing campaign is about one and a half orders of magnitude lower 
than our best limit in power units, though again probing a significantly different redshift. 
The MWA analysis is rapidly improving as we uncover and mitigate systematics to realize
the full sensitivity of the telescope.
The red lines in Figure~\ref{fig:1d_ps} indicate the noise level possible 
to reach with these data if systematics can be overcome.
